---
title: "Statistical Inference Course Project - statinference-011"
author: "Esmit Pérez"
date: "February 16, 2015"
output: html_document
---

#Part 1
##Overview

This is the report for the project for the statistical inference class. In it, I will use simulation to explore inference and do some simple inferential data analysis. The project consists of two parts:

1. A simulation exercise.
2. Basic inferential data analysis.

The format and formulas here included are based off the [outline of the project](https://class.coursera.org/statinference-011/human_grading/view/courses/973519/assessments/4/submissions).

##Simulations

```{r}
library(ggplot2)
lambda <- 0.2
miu <- 1/lambda
stdDev <- 1/lambda
numExponentials <- 40
numSimulations <- 1:1000

# we're dealing with random data, so always set seed to make it reproducible.
set.seed(909)

# obtains the mean of running rexp with 40 exponentials and given lambda
cfunc <- function(v) {mean(rexp(numExponentials, lambda))}

#  for each entry in array of size 1000, run the function
mns = NULL
for (i in 1 : 1000) mns = c(mns, mean(cfunc()))

dat <- data.frame(x = mns)

```


##Sample Mean versus Theoretical Mean

Theoretical mean is 1/lambda:
```{r}
miu
```

The sample mean is:
```{r}
mean(dat$x)
```

##Sample Variance versus Theoretical Variance

Theoretical variance is miu/sqrt(n):
```{r}
miu/sqrt(numExponentials)
```

The sample variance is:
```{r}
var(dat$x)
```

##Distribution

The accompanying graph shows...

```{r}
g <- ggplot(dat, aes(x = x)) + geom_histogram(binwidth=.3, colour = "black", fill="white", aes(y = ..density..)) 
g <- g + stat_function(fun = dnorm, arg=list( mean= miu, sd=sd(dat$x) ), color="red", size=1)

g
```

#Part 2
Now in the second portion of the class, we're going to analyze the ToothGrowth data in the R datasets package. 
Load the ToothGrowth data and perform some basic exploratory data analyses 
```{r}
data(ToothGrowth)
```

Provide a basic summary of the data.
```{r}
str(ToothGrowth)
summary(ToothGrowth)
head(ToothGrowth)
```

Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose. (Only use the techniques from class, even if there's other approaches worth considering)

State your conclusions and the assumptions needed for your conclusions. 