---
title: "Statistical Inference (statinference-011) Course Project - Part 2"
author: "Esmit Pérez"
date: "February 16, 2015"
output: pdf_document
---

```{r setoptions, echo = FALSE}
library(knitr)
# always echo.
opts_chunk$set(echo = TRUE, fig.width = 12, fig.height = 5)
```

#Part 2
Now in the second portion of the class, we're going to analyze the ToothGrowth data in the R datasets package. 

Load the ToothGrowth data and perform some basic exploratory data analyses.

```{r load-tooth-growth}
#load data set from R's provided datasets
data(ToothGrowth)
```


###Provide a basic summary of the data. 

_Note_: Please refer to appendix for a complete summary.

The following plot shows the growth length, per supplement at varying dosages (see R code and graph in appendix, fig 2.1):
```{r, tooth-growth-boxplot, echo=FALSE}
library(ggplot2)
#plot and customize graph
g <- ggplot(ToothGrowth,aes(x = dose, y = len, color=supp))
g <- g + ggtitle("The Effect of Vitamin C on Tooth Growth in Guinea Pigs") 
g <- g + xlab("Dose level (in mg)") + ylab("Length")
g <- g + scale_color_discrete(name="Delivery Method (Supplement)",
                         breaks=c("OJ", "VC"),
                         labels=c("Orange Juice", "Ascorbic Acide"))
g <- g + geom_boxplot()
g <- g + facet_grid(. ~ dose , labeller = function(variable, value){ paste0(value," mg") })
g
```

###Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose. (Only use the techniques from class, even if there's other approaches worth considering)

$H_{0}$ : Supplement _OJ (Orange Juice)_ leads to more growth.


###State your conclusions and the assumptions needed for your conclusions. 

Based on (Week 3 video) (https://class.coursera.org/statinference-011/lecture/243) we can perform a `t.test` to determine whether this is true, assuming variances are equal (see complete output in appendix, fig 2.2):

```{r ttestequal, echo=FALSE}
# Assume variances ARE equal
rbind(t.test(len ~ supp, paired=FALSE, var.equal=T, data=ToothGrowth)$conf,
      t.test(len ~ supp, paired=FALSE, var.equal=T, data=ToothGrowth)$conf
      )
```

Factors in `supp` show `OJ` comes first (see appendix, fig 2.3) and in both tests the result, is the same, column `1`, correspoding to `OJ` (_Orange Juice_) is below zero, meaning it does not lead to more length.


\newpage

#Appendix

## Part 2 - Supporting figures and exploratory analysis, ToothGrowth data.

##Description

"The Effect of Vitamin C on Tooth Growth in Guinea Pigs"

The response is the length of odontoblasts (teeth) in each of 10 guinea pigs at each of three dose levels of Vitamin C (0.5, 1, and 2 mg) with each of two delivery methods (orange juice or ascorbic acid).

##Summary information for `ToothGrowth`:
```{r tooth-growth-summary}
str(ToothGrowth)
summary(ToothGrowth)
head(ToothGrowth)
```

##Example `coplot` as per `help(ToothGrowth)`:

```{r tooth-growth-coplot}
require(graphics)
coplot(len ~ dose | supp, data = ToothGrowth, panel = panel.smooth,
       xlab = "ToothGrowth data: length vs dose, given type of supplement")
```


### Figure 2.1
```{r, tooth-growth-boxplot-app}
#plot and customize graph
g <- ggplot(ToothGrowth,aes(x = dose, y = len, color=supp))
g <- g + ggtitle("The Effect of Vitamin C on Tooth Growth in Guinea Pigs") 
g <- g + xlab("Dose level (in mg)") + ylab("Length")
g <- g + scale_color_discrete(name="Delivery Method (Supplement)",
                         breaks=c("OJ", "VC"),
                         labels=c("Orange Juice", "Ascorbic Acide"))
g <- g + geom_boxplot()
g <- g + facet_grid(. ~ dose , labeller = function(variable, value){ paste0(value," mg") })
g
```

###Figure 2.2
```{r ttestequal-app}
# Assume variances ARE equal
rbind(t.test(len ~ supp, paired=FALSE, var.equal=T, data=ToothGrowth)$conf,
      t.test(len ~ supp, paired=FALSE, var.equal=T, data=ToothGrowth)$conf
      )
```

####Figure 2.3
```{r factorsofsupp-app}
factor(ToothGrowth$supp)
```